The Spatial and Temporal Domains of Modern Ecology

### 2017-06-20 Rmd/space-time.Rmd
Full revision, including discussion and NatEcolEvol style abstract.

p_root <- set_base_path("ecoscales")
p_dat <- fp(p_root, "external/data/result")
p_calib <- fp(p_root, "external/data/calibration")
p_fig <- fp(p_root, "paper/figures")

@@

```{r, eval=FALSE}
# p <- c(0.64, 0.84, 0.53, 1.21)  # from calibration.Rmd
# p <- c(0.58, 0.82, 0.99, 1.24)  # from calibration.Rmd
##     plot_res   n_sites   act_ext   eff_ext t_btwn_samp  act_dur   eff_dur
## 1: 0.5769693 0.3821107 0.7196559 0.4049972     1.01134 1.253045 0.5284913
##    plot_res n_sites  act_ext  eff_ext t_btwn_samp  act_dur  eff_dur
##       0.494   0.457    0.749    0.499       0.475    1.093    0.338
##    plot_res n_sites  act_ext  eff_ext t_btwn_samp  act_dur  eff_dur
##       0.577   0.382     0.72    0.405       1.045    1.261    0.643
p <- c(0.58, 0.72, 0.41, 1.05, 1.26, 0.64)  # from calibration.Rmd

@@

# Run bootstrap
set.seed(1)
bootperturb <- resamp_func(datf, p = p, iter = 1000) 
# bootperturb2 <- resamp_func(datf, p = p, iter = 1000)  # 10% perturb
# all(bootperturb[[200]]$plot_res == bootperturb2[[200]]$plot_res)
# rm(bootperturb2)
```

@@

# spatial
# plot(1:10, xlab = expression(paste("mm"^2)), xaxt = "n")
# axis(1, at = 1:8, labels = lab, las = 2)

@@

### Figure 1

Histograms for each of the four main dimensions.
```{r, eval=FALSE}
# length(which(datf$study_type %in% c("remote", "other"))) / nrow(datf)
datf[t_btwn_samp == 0, .N] / datf[, .N]  # 37% of samples are once-offs
datf[act_dur == samp_dur, .N] / nrow(datf)
datf[(act_dur == samp_dur) & t_btwn_samp != 0]
datf[(act_dur != samp_dur) & t_btwn_samp == 0]

@@

# length(which(hdat$res >= -3 & hdat$res <= 0)) / length(hdat$res)
# length(which(hdat$res <= 4)) / length(hdat$res)
# sum(hdatstat[[1]]$counts[4:6]) / sum(hdatstat[[1]]$counts)

@@

# interval
# hdatst <- hdatstat[[3]]$density
hdatst <- hdatstat_bt$int$mu
l <- length(hdatst)
hds <- c("norep", "<daily", "<weekly",  "day-mo", "mo-yr", "yr-dec")
inds <- list(l, 1:3, 1:4, 4:5, 6, 7:8)
sapply(1:length(inds), function(x) {
  v <- round(sum(hdatst[inds[[x]]]), 1)
  names(v) <- hds[x]
  v
})

@@

# datf[, plot(log10(n_sites), log10(plot_res))]
# dat[like(study_type, "field"), plot(log10(n_sites), log10(plot_res))]
# datf[like(study_type, "field"), plot(log10(plot_res), log10(act_ext), 
#                                      ylim = c(-10, 10))]
# datf[like(study_type, "remote"), 
#      points(log10(plot_res), log10(act_ext), pch = 20, col = "red")]
# datf[like(study_type, "automated"), 
#      points(log10(plot_res), log10(act_ext), pch = 20, col = "blue")]
# lines(aaxis1$logres, aaxis2$logarea[-c(16:17)] )
# datf[, .N / nrow(datf) * 100, by = st]

# datf[, plot(log10(t_btwn_samp), log10(act_dur), pch = 20)]
# datf[which(hdat[[3]] < -1.380211), 
#      points(log10(t_btwn_samp), log10(act_dur), pch = 20, col = "blue")]
# datf[like(study_type, "automated"), 
#      points(log10(t_btwn_samp), log10(act_dur), pch = 20, col = "red")]

@@

# pdf("paper/figures/res_v_extent.pdf", width = 7, height = 2.5)
pdf("paper/figures/kde45.pdf", width = 7, height = 6)
par(mfrow = c(2, 2), mar = c(6, 6, 0.3, 3), oma = c(0, 0, 0.5, 1), 
    mgp = c(2, 0.25, 0))
# resolutions
# image(rres3, col = rev(terrain.colors(30)), breaks = brks,
image(rres3, axes = FALSE, xlab = "", ylab = "", #rev(terrain.colors(cuts))) 
      col = cols)
axis(1, at = taxis1$logdays[-c(5)], labels = tlab1[-c(5)], 
     las = 2, tcl = -0.2, cex.axis = cxa)
axis(2, at = aaxis1$logres[-1], labels = alab1[-1], las = 2, tcl = -0.2, 
     cex.axis = cxa)
mtext("Interval", 1, cex = cxa, line = ll)
mtext("Resolution", 2, cex = cxa, line = ll)
brks <- brkfun(rres3, 0.01, cuts)  # breaks
aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
              cex.axis = cxa, tcl = -0.1)
plot(rres3, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     #legend.shrink = 0.9, breaks = brks, col = rev(terrain.colors(30)))
     legend.shrink = 0.9, col = cols)#inferno(cuts))
     #col = rev(terrain.colors(cuts)))
mtext("A", side = 3, line = -1, cex = 0.8, adj = 0.05, col = "white")
# for(i in 1:length(stps)) {
#   points(resdat3$x[stps[[i]]], resdat3$y[stps[[i]]], pch=pchs[[i]], cex=cexs[i])
# }

# extents
# image(rext3, col = rev(terrain.colors(30)), axes = FALSE, xlab = "", 
image(rext3, axes = FALSE, xlab = "", ylab = "", col = cols)
axis(1, at = taxis2$logdays[-5], labels = tlab2[-5], las = 2,
     cex.axis = cxa, tcl = -0.2)
axis(2, at = aaxis2$logarea, labels = alab2, las = 2, cex.axis = cxa,
     tcl = -0.2)
brks <- brkfun(rext3, 0.01, cuts)  # breaks
aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
              cex.axis = cxa, tcl = -0.1)
mtext("Duration", 1, cex = cxa, line = ll)
mtext("Extent", 2, cex = cxa, line = ll)
plot(rext3, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     legend.shrink = 0.9, col = cols)
mtext("B", side = 3, line = -1, cex = 0.8, adj = 0.05, col = "white")

# spatial
image(sres, col = cols, axes = FALSE, xlab = "", ylab = "") 
axis(1, at = aaxis1$logres, labels = alab1, las = 2, tcl = -0.2, cex.axis = cxa)
axis(2, at = aaxis2$logarea, labels = alab2, las = 2, tcl = -0.2, 
     cex.axis = cxa)
mtext("Resolution", 1, cex = cxa, line = ll)
mtext("Extent", 2, cex = cxa, line = ll)
brks <- brkfun(sres, 0.01, cuts)  # breaks
aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
              cex.axis = cxa, tcl = -0.1)
plot(sres, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     legend.shrink = 0.9, col = cols)
mtext("C", side = 3, line = -1, cex = 0.8, adj = 0.05, col = "white")
# for(i in 1:length(stps)) {
#   points(spatdat$x[stps[[i]]], spatdat$y[stps[[i]]], pch=pchs[[i]], cex=cexs[i])
# }

@@

@@

### Figure 3
```{r, eval=FALSE}
# actual versus effective duration
# kdat <- bootp_dt  # bootstrap
# # tempdat2 <- data.table(x = log10(kdat$act_dur), y = log10(kdat$eff_dur))
# tempdat2 <- data.table(x = log10(kdat$eff_dur), y = log10(kdat$act_dur))
# dinds <- kdat[, which(t_btwn_samp == 365 * 10000)]  # drop unreplicated obs
# tempdat2 <- tempdat2[-dinds, ]
# tempdat2 <- tempdat2[!is.na(x) & !is.na(y)] # cut out null values
# tempdat2[, n := 1]  # set up count column
# tx <- range(taxis2$logdays)
# tx <- c(floor(ex[1]), 5.56) ### should this be 5.57?  #ceiling(ex[2]))
# tempdat2[x < tx[1], x := tx[1]]
# tempdat2[x > tx[2], x := tx[2]]
# ty <- range(taxis2$logdays)
# ty <- c(floor(ex[1]), 5.56) ### should this be 5.57?  #ceiling(ex[2]))
# tempdat2[y < tx[1], y := tx[1]]
# tempdat2[y > tx[2], y := tx[2]]
# coordinates(tempdat2) <- ~x + y
# 
# # extent
# # extdat4 <- data.table(x = log10(kdat$act_ext), y = log10(kdat$eff_ext))
# extdat4 <- data.table(x = log10(kdat$eff_ext), y = log10(kdat$act_ext))
# extdat4 <- extdat4[!is.na(x) & !is.na(y)] # cut out null values
# extdat4[, n := 1]  # set up count column
# ex <- range(aaxis2$logarea)
# extdat4[x < ex[1], x := ex[1]]
# extdat4[x > ex[2], x := ex[2]]
# ey <- range(aaxis2$logarea)
# extdat4[y < ey[1], y := ey[1]]
# extdat4[y > ey[2], y := ey[2]]
# coordinates(extdat4) <- ~x + y
# 
# # densities
# # temporal 
# tres2 <- kdensity(tx[1], tx[2], ty[1], ty[2], 0.1, tempdat2, 1)
# tres2 <- (tres2 / cellStats(tres2, sum)) * 100
# 
# # extent
# rext4 <- kdensity(ex[1], ex[2], ey[1], ey[2], 0.1, extdat4, 1)
# rext4 <- (rext4 / cellStats(rext4, sum)) * 100
# 
# # cxa = 0.7

# plot params
ex <- range(aaxis2$logarea)
ey <- range(aaxis2$logarea)
ll <- 3.5
cxa = 0.7
stps <- sapply(c(3, 4), function(x) which(kdat$st == x))  # IDs rs/auto/paleo
pchs <- list(1, "+")
cexs <- c(0.4, 0.6)
cuts <- 40
bump <- 3
cols <- inferno(cuts + bump)[-c(2:(bump + 1))]
# # plot(1:cuts, col = cols, pch = 20)
# 
# # par(mfrow = c(1, 2))
# # datf[, {
# #   # sel = which()
# #   eext = log10(eff_ext)
# #   aext = log10(act_ext)
# #   edur = log10(eff_dur[t_btwn_samp != 365 * 10000])
# #   adur = log10(act_dur[t_btwn_samp != 365 * 10000])
# #   plot(aext, eext - aext)
# #   plot(adur, edur - adur)
# #   length(edur)
# # }]

# Difference in effective versus actual extent/duration against actual ext/dur
hdat_bs2 <- lapply(1:length(bootperturb), function(x) {  # x <- 1
  print(x)
  DT <- bootperturb[[x]]
  edb <- DT[, lapply(.SD, log10), .SDcol = c("eff_ext", "act_ext")]
  # edb <- edb[eff_ext >= act_ext, ]  # drop eff_dur < act
  setnames(edb, c("eff_ext", "act_ext"), c("eext", "aext"))
  edb[, dif := eext - aext]

  tdb <- DT[t_btwn_samp != 365 * 10000, lapply(.SD, log10), 
            .SDcol = c("eff_dur", "act_dur")]
  # tdb <- tdb[eff_dur >= act_dur, ]  # drop eff_dur < act
  setnames(tdb, c("eff_dur", "act_dur"), c("edur", "adur"))
  # print(paste(edb[, .N], tdb[, .N]))
  tdb[, dif := edur - adur]
  return(list("edb" = edb, "tdb" = tdb, #"elm" = elm, "tlm" = tlm, 
              "Ns" = c(edb[, .N], tdb[, .N])))
})

@@

# # Kernel density version
# pdf("paper/figures/act_v_eff.pdf", width = 7, height = 3)
# par(mfrow = c(1, 2), mar = c(4.5, 4.5, 0.3, 3), oma = c(0, 0, 0.2, 1), 
#     mgp = c(2, 0.25, 0))
# 
# mtl <- 3
# # spatial extents
# image(rext4, axes = FALSE, xlab = "", ylab = "", col = cols)
# axis(1, at = aaxis2$logarea, labels = alab2, las = 2,
#      cex.axis = cxa, tcl = -0.2)
# axis(2, at = aaxis2$logarea, labels = alab2, las = 2, cex.axis = cxa,
#      tcl = -0.2)
# brks <- brkfun(rext4, 0.01, cuts)  # breaks
# aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
#               cex.axis = cxa, tcl = -0.1)
# mtext("Actual extent", 2, cex = cxa, line = mtl)
# mtext("Effective extent", 1, cex = cxa, line = mtl)
# plot(rext4, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
#      legend.shrink = 0.9, col = cols)
# mtext("A", side = 3, line = -1, cex = 0.8, adj = 0.05, col = "white")
# 
# # temporal extents
# image(tres2, axes = FALSE, xlab = "", ylab = "", col = cols)
# axis(1, at = taxis1$logdays[-c(5)], labels = tlab1[-c(5)], 
#      las = 2, tcl = -0.2, cex.axis = cxa)
# axis(2, at = taxis1$logdays[-c(5)], labels = tlab1[-c(5)], las = 2, tcl = -0.2, 
#      cex.axis = cxa)
# mtext("Actual duration", 2, cex = cxa, line = mtl)
# mtext("Effective duration", 1, cex = cxa, line = mtl)
# brks <- brkfun(tres2, 0.01, cuts)  # breaks
# aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
#               cex.axis = cxa, tcl = -0.1)
# plot(tres2, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
#      legend.shrink = 0.9, col = cols)
#      #col = rev(terrain.colors(cuts)))
# mtext("B", side = 3, line = -1, cex = 0.8, adj = 0.05, col = "white")
# dev.off()
```

## Supplemental Analyses

@@

remote <- datf[study_type == "remote", eff_ext]
notremote <- datf[study_type != "remote", eff_ext]
# a <- (length(remote) + rsinc) / (nrow(datf))
# b <- length(remote) / nrow(datf)
# a / b
# ((length(remote) + rsinc) / (length(notremote)) / length(b))

# (rsext * rs2017 + nonrsext * (1 - rs2017)) / 
#   (rsext * rs2014 + nonrsext * (1 - rs2014))

@@

>>>>

p_root <- set_base_path("ecoscales")
p_dat <- fp(p_root, "external/data/result")
p_calib <- fp(p_root, "external/data/calibration")
p_fig <- fp(p_root, "paper/figures")
p_til <- fp(p_root, "external/data/misc")

@@

```{r, eval=FALSE}
p <- c(0.58, 0.72, 0.41, 1.05, 1.26, 0.64)  # from calibration.Rmd

@@

# Run bootstrap
set.seed(1)
bootperturb <- resamp_func(datf, p = p, iter = 1000) 
```

@@

@@

### Figure 1 
Histograms for each of the four main dimensions.
```{r, eval=FALSE}
# length(which(datf$study_type %in% c("remote", "other"))) / nrow(datf)
datf[t_btwn_samp == 0, .N] / datf[, .N]  # 37% of samples are once-offs
datf[act_dur == samp_dur, .N] / nrow(datf)
datf[(act_dur == samp_dur) & t_btwn_samp != 0]
datf[(act_dur != samp_dur) & t_btwn_samp == 0]

@@

@@

# interval
# hdatst <- hdatstat[[3]]$density
hdatst <- hdatstat_bt$int$mu
l <- length(hdatst)
hds <- c("norep", "<daily", "<weekly", "<minute", "min-hour", 
         "hr-day", "day-wk", "wk-mo", "mo-yr", "yr-dec", "dec-cen", "cen-mil")
inds <- list(l, 1:3, 1:4, 1, 2, 3, 4, 5, 6, 7, 8, 9)
sapply(1:length(inds), function(x) {
  v <- round(sum(hdatst[inds[[x]]]), 1)
  names(v) <- hds[x]
  v
})

@@

@@

# pdf("paper/figures/res_v_extent.pdf", width = 7, height = 2.5)
pdf("paper/figures/kde45.pdf", width = 7, height = 6)
par(mfrow = c(2, 2), mar = c(6, 6, 0.3, 3), oma = c(0, 0, 0.5, 1), 
    mgp = c(2, 0.25, 0))
# resolutions
# image(rres3, col = rev(terrain.colors(30)), breaks = brks,
image(rres3, axes = FALSE, xlab = "", ylab = "", #rev(terrain.colors(cuts))) 
      col = cols)
axis(1, at = taxis1$logdays[-c(5)], labels = tlab1[-c(5)], 
     las = 2, tcl = -0.2, cex.axis = cxa)
axis(2, at = aaxis1$logres[-1], labels = alab1[-1], las = 2, tcl = -0.2, 
     cex.axis = cxa)
mtext("Interval", 1, cex = cxa, line = ll)
mtext("Resolution", 2, cex = cxa, line = ll)
brks <- brkfun(rres3, 0.01, cuts)  # breaks
aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
              cex.axis = cxa, tcl = -0.1)
plot(rres3, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     #legend.shrink = 0.9, breaks = brks, col = rev(terrain.colors(30)))
     legend.shrink = 0.9, col = cols)#inferno(cuts))
     #col = rev(terrain.colors(cuts)))
mtext("A", side = 3, line = -1, cex = 0.8, adj = 0.05, col = "white")

# extents
image(rext3, axes = FALSE, xlab = "", ylab = "", col = cols)
axis(1, at = taxis2$logdays[-5], labels = tlab2[-5], las = 2,
     cex.axis = cxa, tcl = -0.2)
axis(2, at = aaxis2$logarea, labels = alab2, las = 2, cex.axis = cxa,
     tcl = -0.2)
brks <- brkfun(rext3, 0.01, cuts)  # breaks
aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
              cex.axis = cxa, tcl = -0.1)
mtext("Duration", 1, cex = cxa, line = ll)
mtext("Extent", 2, cex = cxa, line = ll)
plot(rext3, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     legend.shrink = 0.9, col = cols)
mtext("B", side = 3, line = -1, cex = 0.8, adj = 0.05, col = "white")

# spatial
image(sres, col = cols, axes = FALSE, xlab = "", ylab = "") 
axis(1, at = aaxis1$logres, labels = alab1, las = 2, tcl = -0.2, cex.axis = cxa)
axis(2, at = aaxis2$logarea, labels = alab2, las = 2, tcl = -0.2, 
     cex.axis = cxa)
mtext("Resolution", 1, cex = cxa, line = ll)
mtext("Extent", 2, cex = cxa, line = ll)
brks <- brkfun(sres, 0.01, cuts)  # breaks
aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
              cex.axis = cxa, tcl = -0.1)
plot(sres, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
     legend.shrink = 0.9, col = cols)
mtext("C", side = 3, line = -1, cex = 0.8, adj = 0.05, col = "white")

@@

save(bootp_dt, file = fp(p_dat, "derived/bootresults.rda"))
```

@@

### Figure 3

Actual versus effective duration/extent
```{r, eval=FALSE}

# plot params
ex <- range(aaxis2$logarea)
ey <- range(aaxis2$logarea)
ll <- 3.5
cxa = 0.7
stps <- sapply(c(3, 4), function(x) which(kdat$st == x))  # IDs rs/auto/paleo
pchs <- list(1, "+")
cexs <- c(0.4, 0.6)
cuts <- 40
bump <- 3
cols <- inferno(cuts + bump)[-c(2:(bump + 1))]

# Difference in effective versus actual extent/duration against actual ext/dur
hdat_bs2 <- lapply(1:length(bootperturb), function(x) {  # x <- 1
  print(x)
  DT <- bootperturb[[x]]

  # extent measures log10 transformed
  edb <- DT[, lapply(.SD, log10), .SDcol = c("eff_ext", "act_ext")]
  setnames(edb, c("eff_ext", "act_ext"), c("eext", "aext"))
  edb[, dif := eext - aext]  # difference between extent and actual extent

  # duration measures log10 transformed, unreplicated dropped
  tdb <- DT[t_btwn_samp != 365 * 10000, lapply(.SD, log10), 
            .SDcol = c("eff_dur", "act_dur")]
  setnames(tdb, c("eff_dur", "act_dur"), c("edur", "adur"))
  # print(paste(edb[, .N], tdb[, .N]))
  tdb[, dif := edur - adur]  # difference between duration and actual duration
  return(list("edb" = edb, "tdb" = tdb, #"elm" = elm, "tlm" = tlm, 
              "Ns" = c(edb[, .N], tdb[, .N])))
})

@@

## Additional Analyses

@@

remote <- datf[study_type == "remote", eff_ext]
notremote <- datf[study_type != "remote", eff_ext]

@@

### Scales in earlier studies
#### Duration from Tilman, 1989
```{r, echo=FALSE}
load(fp(p_dat, "derived/bootresults.rda"))
```
```{r}
std <- function(x) (x - min(x)) / diff(range(x))

# Tilman (1989), Figure 6.1 (visually interpreted)
# Data extracted from figure using http://arohatgi.info/WebPlotDigitizer/app/
tilman <- fread(fp(p_til, "tilman1989.csv"))

# Find n and proportion of non-experimental observations per year
# find mean/median value of data > 100 years in our observed data
gt100 <- bootp_dt[eff_dur > (365 * 100), {
  list("med" = median(eff_dur) / 365, "mu" = mean(eff_dur) / 365) 
}]

# find weights and proportion/year
tilman_noexp <- tilman[, {
  a = round(623 * prop[fig == 1] * (1 / sum(prop[fig == 1])))
  b = round(180 * prop[fig == 2] * (1 / sum(prop[fig == 2])))
  padj = a - b
  w = padj / sum(padj)
  list("yrmu" = yrmu[fig == 1], "n" = padj, "prop" = w)
}]

# median and mean per year, assuming 150 years for studies > 100
tilmedmu <- tilman_noexp[, rep(yrmu, n), by = n][, {
  list("med" = median(V1), "mean" = mean(V1))
}]

# test additional upper bin years
gt1002 <- unname(c(200, 500, 1000, unlist(gt100)))
tilmedmu2 <- rbindlist(lapply(gt1002, function(x) {
  DT2 = copy(tilman_noexp)[yrmu == 150, yrmu := x]
  DT2[, rep(yrmu, n), by = n][, list("med" = median(V1), "mean" = mean(V1))]
}))

durmedmu <- bootp_dt[, {
  list("med" = median(eff_dur[t_btwn_samp != (365 * 10000)], na.rm = TRUE) / 365, 
       "mean" = mean(eff_dur[t_btwn_samp != (365 * 10000)], na.rm = TRUE) / 365, 
       "medall" = median(eff_dur, na.rm = TRUE) / 365, 
       "meanall" = mean(eff_dur, na.rm = TRUE) / 365)
}]

# And calculate only for years less than 100
tilmedmu_lt100 <- tilman_noexp[yrmu < 150, rep(yrmu, n), by = n][, {
  list("med" = median(V1), "mean" = mean(V1))
}]

durmedmu_lt100 <- bootp_dt[eff_dur < (365 * 100), {
  list("med" = median(eff_dur[t_btwn_samp != (365 * 10000)], na.rm = TRUE) / 365, 
       "mean" = mean(eff_dur[t_btwn_samp != (365 * 10000)], na.rm = TRUE) / 365, 
       "medall" = median(eff_dur, na.rm = TRUE) / 365, 
       "meanall" = mean(eff_dur, na.rm = TRUE) / 365)
}]

Median, mean duration from Tilman (1989)
```{r}
round(rbind(tilmedmu, tilmedmu2), 2)
```

Tilman with studies > 100 years dropped
```{r}
round(tilmedmu_lt100, 2)
```

Tilman (1989) duration of observational studies by bin
```{r}
tilman_noexp[, {
  a = round(prop * 100, 1)
  names(a) = tilman[fig == 1, yr]
  rbind(a, n)
}]
```

Median, mean duration from this study
```{r}
round(durmedmu, 2)
```

This study with years > 100 dropped
```{r}
round(durmedmu_lt100, 2)
```

#### Kareiva and Anderson (1989)

Resolution versus this study
```{r}
# helper func
q80 <- function(x) quantile(x, 0.8, na.rm = TRUE)
q90 <- function(x) quantile(x, 0.9, na.rm = TRUE)
q95 <- function(x) quantile(x, 0.95, na.rm = TRUE)

kareiva <- fread(fp(p_til, "kareiva_anderson1989.csv"))
a <- kareiva[, {
  list("med" = median(res^2), "mean" = mean(res^2), "80" = q80(res^2), 
       "90" = q90(res^2), "95" = q95(res^2))
}]
b <- bootp_dt[, {
  list("med" = median(plot_res), "mean" = mean(plot_res), "80" = q80(plot_res), 
       "90" = q90(plot_res), "95" = q95(plot_res))
}]
study <- c("Kareiva & Anderson, 1989", "This study")
cbind(study, round(rbind(a, b), 2))
```

#### Porter et al (2005)
Interval
```{r}
porter <- fread(fp(p_til, "porter.csv"))
porter <- porter[sample == 1]
porter[, ext := ext^2 / 10000]

study <- c("Porter et al, 1989", "This study")
a <- porter[int > 0, {
  list("med" = median(int), "mean" = mean(int), "80" = q80(int), "90" = q90(int),
       "95" = q95(int))
}]

b <- bootp_dt[t_btwn_samp != (365 * 10000), {
  list("med" = median(t_btwn_samp), "mean" = mean(t_btwn_samp), 
       "80" = q80(t_btwn_samp), "90" = q90(t_btwn_samp), "95" = q95(t_btwn_samp))
}]
cbind(study, round(rbind(a, b)))
```

Extent
```{r}
study <- c("Porter et al, 1989", "This study", "Ratio")
a <- porter[, {
  list("med" = median(ext), "mean" = mean(ext), "80" = q80(ext), 
       "90" = q90(ext), "95" = q95(ext))
}] 
b <- bootp_dt[, {
  list("med" = median(eff_ext, na.rm = TRUE), 
       "mean" = mean(eff_ext, na.rm = TRUE), "80" = q80(eff_ext), 
       "90" = q90(eff_ext), "95" = q95(eff_ext))
}]
cbind(study, round(rbind(a, b, b / a)))
```

### Share of remote sensing, automated, and paleo observations
```{r}

aext <- 10^4
datf[, {
  a = length(which(act_ext > aext))
  b1 = length(which((act_ext > aext) & like(study_type, "remote")))
  b2 = length(which((act_ext > aext) & like(study_type, "other")))
  round(c("remote" = b1 / a * 100, "other" = b2 / a * 100), 1)
}]

adur <- 365
datf[, {
  a = length(which(act_dur > adur))
  b1 = length(which((act_dur > adur) & like(study_type, "auto")))
  b2 = length(which((act_dur > adur) & like(study_type, "paleo")))
  round(c("automated" = b1 / a * 100, "paleo" = b2 / a * 100), 1)
}]

datf[, {
  a = length(which(act_dur <= adur))
  b1 = length(which((act_dur <= adur) & like(study_type, "auto")))
  b2 = length(which((act_dur <= adur) & like(study_type, "paleo")))
  round(c("automated" = b1 / a * 100, "paleo" = b2 / a * 100), 1)
}]

### 2017-06-20 paper/SI.tex
Everything through final discussion section revised
Full revision, including discussion and NatEcolEvol style abstract.

@@

\begin{table}[ht]
\captionsetup{width=1\linewidth}
\caption{Estimates of variability calculated from each observer's estimates of the spatial and temporal scales of ecological observations reported within a common set of the 20 papers used for calibration. Variability is expressed as the coefficient of variation (CV; standard deviation divided by mean multiplied by 100) between each observer's overall mean, and as the mean CV of observers' estimates for individual records. }
\vspace{-4 pt}
\begin{center}
\begin{tabular}{rR{1.5cm}R{1.5cm}R{1.5cm}R{1.5cm}R{1.5cm}R{1.5cm}}
\hline
& \multicolumn{3}{c}{\textbf{Spatial}} & \multicolumn{3}{c}{\textbf{Temporal}}\\\cline{2-7}
%Value & Resolution & Extent & \multirow{2}{*}{Actual} & Interval & Duration & Actual Duration \\
Value & Resolution & Extent & Actual Extent & Interval & Duration & Actual Duration \\
\hline
CV of overall mean & 50 & 50 & 75 & 48 & 34 & 109 \\
Mean of record-wise CV & 57 & 41 & 72 & 104 & 64 & 124 \\
\hline
\end{tabular}
\end{center}
\label{default}
\end{table}%

>>>>

\subsubsection*{Extracting and analyzing data from earlier meta-analyses}

419 observational studies lasting $\leq100 years; longer studies were excluded because the upper duration bound was not published)

reviewed in this study

@@

\begin{table}[ht]
\captionsetup{width=1\linewidth}
\caption{Estimates of variability calculated from each observer's estimates of the spatial and temporal scales of ecological observations reported within a common set of the 20 papers used for calibration. Variability is expressed as the coefficient of variation (CV; standard deviation divided by mean multiplied by 100) between each observer's overall mean, and as the mean CV of observers' estimates for individual records. }
\vspace{-4 pt}
\begin{center}
\begin{tabular}{rR{1.5cm}R{1.5cm}R{1.5cm}R{1.5cm}R{1.5cm}R{1.5cm}}
\hline
& \multicolumn{3}{c}{\textbf{Spatial}} & \multicolumn{3}{c}{\textbf{Temporal}}\\\cline{2-7}
%Value & Resolution & Extent & \multirow{2}{*}{Actual} & Interval & Duration & Actual Duration \\
Value & Resolution & Extent & Actual Extent & Interval & Duration & Actual Duration \\
\hline
CV of overall mean & 50 & 50 & 75 & 48 & 34 & 109 \\
Mean of record-wise CV & 57 & 41 & 72 & 104 & 64 & 126 \\
\hline
\end{tabular}
\end{center}
\label{default}
\end{table}%
